Very simplified likelihood

### 2023-11-22 inst/very_simplified_likelihood.Rmd
kickoff for simplified likelihood

```{r tryit}
small <- 1e-3
f_gamma <- function(a,r,n,log=FALSE){ # vectorised in 'a', but screws up for b=1/integer
  b <- (1-a)/n   # 0 <= b <= 1/n
  out <- log(a) + lgamma(1-1/b) - lgamma(1+r-1/b)
  if(!log){out <- exp(out)}
  return(out)
}

f_single_direct <- function(a,r,n,log=FALSE){ # not vectorised in anything
  stopifnot(length(a)==1)
  stopifnot(length(r)==1)
  stopifnot(length(n)==1)
  b <- (1-a)/n   # 0 <= b <= 1/n
  if(log){
      return(log(a)+r*log(b)-sum(log(1-seq_len(r)*b)))
     } else {
      return(a * b^r / prod(1-seq_len(r)*b))
     }
}

f <- function(a,r,n,log=FALSE){  # vectorized in a
out <- a*0
if(log){
    for(rr in r){  out <- out + f_single_direct(a,rr,n,log=TRUE)}
  } else {
    out[] <- 1
    for(rr in r){  out <- out * f_single_direct(a,rr,n,log=FALSE)}
  }
  return(out)
}

n <- 10
a <- seq(from=small,to=1-small,len=100)
plot(a,f(a,r=3,n=n))
L <- f(a,r=3,n=n,log=TRUE)
L <- L-max(L)
plot(a,L,ylim=c(-5,0))
f(0.9,3,10)
```

>>>>

```{r tryit}
small <- 1e-3

f_single_direct <- function(a,r,n,log=FALSE){  # not vectorised in anything
  b <- (1-a)/n   # 0 <= b <= 1/n
  if(log){
      return(log(1-n*b)+r*log(b)-sum(log(1-seq_len(r)*b)))
     } else {
      return((1-n*b)*b^r-prod(1-(1:3)*b))
     }
}

f <- function(a,r,n,log=FALSE){
out <- a*0
if(log){
    for(rr in r){  out <- out + f_single(a,rr,n,log=TRUE)}
  } else {
    out[] <- 1
    for(rr in r){  out <- out * f_single(a,rr,n,log=FALSE)}
  }
  return(out)
}

f_gamma <- function(a,r,n,log=FALSE){ # vectorised in 'a', but screws up for b=1/integer
  b <- (1-a)/n   # 0 <= b <= 1/n
  out <- log(a) + lgamma(1-1/b) - lgamma(1+r-1/b)
  if(!log){out <- exp(out)}
  return(out)
}
```

Now use it:

```{r}
 a <- seq(from=0.01,to=0.99,by=0.01)
 L <- f(a,r=c(3,4,5),10)
plot(a,L-max(L))
```
